---
title: "social_network_analysis"
author: "Fan Ouyang"
date: "August 22, 2017"
output: html_document
---
## An exampler of social network analysis with two different formats
This example shows how to use popular SNA packages (i.e., *sna* and *igraph*) to conduct network analysis for are two primary data formats, namely adjacency matrix and edgelist.

### 1.adjacency matrix (wide-format)
The first format is adjacency matrix (refer to http://faculty.ucr.edu/~hanneman/nettext/C5_%20Matrices.html for more details). This format can be considered as wide-format (in contrast with the second format: long-format)
#### one-mode network analysis
Adjacency matrix dataset is usually used in one-mode network analysis (e.g. learner-learner interaction network). Pls refer to data folder in my github repo for the csv file. In this example, I use *edge width* to show degree(indegree and outdegree), and *color* to show students and the instructor. Here I use *sna* package to show how to plot socio-graph.
 
```{r message = FALSE, warning = FALSE}
#don't forget to set working directory at the beginning setwd("/Users/lt/desktop")
require(sna)
# read data in
one_mode_mat<- as.matrix(read.csv("one_mode_ad_mat.csv", row.names=1))
colnames(one_mode_mat)=c("A","B","C","Ins")
rownames(one_mode_mat)=c("A","B","C","Ins")
one_mode_mat
# calculate degree
id<-degree(one_mode_mat,gmode="digraph",cmode="indegree")
od<-degree(one_mode_mat,gmode="digraph",cmode="outdegree")
# create color vector to show different colors for student and instructor
stu <- rep("magenta3", 3)
ins <- rep("darkolivegreen4", 1) # final one is the instructor
color <- c(stu, ins)
# plot one-mode network graph
gplot(one_mode_mat, gmode="graph", displaylabels=TRUE, 
      label.cex = 0.7, label.col = "gray8", label.pos = 5,
      vertex.cex=2, vertex.col = color, 
      edge.col = "gray38", edge.lwd=(id+od)/3)
```

#### two-mode network analysis
Adjacency matrix can also be used in two-mode network analysis. Here is another example of using adjacency matrix to create learner-activity participation network. Pls refer to data folder in my github repo for the csv file. In this example, I use *color* and *shape* to separately show people and activities and *edge width* to show a person's participation frequency in an activity. Here I use *sna* package to show how to plot socio-graph.

```{r message = FALSE, warning = FALSE}
require(sna)
two_mode_mat<- as.matrix(read.csv("two_mode_ad_mat.csv",row.names=1))
# create edge vector, saving participation frequency
edge<-c(two_mode_mat[,1],two_mode_mat[,2],two_mode_mat[,3],two_mode_mat[,4])
# create color vector to show different colors for people and activities
people <- rep("magenta3", 4)
activity <- rep("darkolivegreen4", 4)
color <- c(people, activity)
# plot two-mode network graph
gplot(two_mode_mat, gmode = "twomode", usearrows = F, displaylabels = T, 
      label.cex = 0.7, label.col = "gray8", label.pos = 5, 
      vertex.col = color, vertex.cex = 2,
      edge.col = "gray38", edge.lwd=edge/2,
      mode="fruchtermanreingold")
```

### 2.edgelist format (long-format)
#### one-mode(or two-mode) network analysis
Like the way adjacency matrix is used for one-mode and two-mode network analysis, edgelist format can also be used for both one-mode and two-mode network analysis. Here I use *igraph* package to show how to plot socio-graph.

```{r message = FALSE, warning = FALSE}
require(igraph)
require(reshape2)
# convert wide-format matrix to long-format edgelist by using melt function
one_mode_ed=melt(one_mode_mat)
# variable names
colnames(one_mode_ed)=c("participant01","participant02","frequency")
# convert edgelist to igraph format
one_mode_graph=graph_from_data_frame(one_mode_ed,directed=TRUE)
# plot with igraph, please refer to http://igraph.org/r/doc/plot.common.html for more details
plot(one_mode_graph,
     vertex.size=5,
     vertex.label.dist=0.5, vertex.color="red",edge.arrow.size = 0.5)
```

### write two functions to calculate node-level and network-level measures using *sna* and *igraph* respectively
```{r}
# use sna package to calculate node-level centralities, make sure input is in network format
node.centrality<-function(x){
           require(sna)
           central.nodes<-cbind(degree(x,cmode="indegree"), degree(x,cmode="outdegree"), betweenness(x,rescale=T),
                                closeness(x,cmode="directed",rescale=T))
           colnames(central.nodes)<-c("indegree","outdegree","betweenness","closeness")
           rownames(central.nodes)<-x%v%"vertex.names"
           
           list(central.nodes)
         }


# use igraph to calculate network-level measures, make sure input is in igraph format
network.measure<-function(x){
      require(igraph)
      graph.measure<-cbind(as.numeric(length(V(x))),
                        as.numeric(length(E(x))),
                        as.numeric(diameter(x)),
                        as.numeric(graph.density(x)),
                        as.numeric(reciprocity(x)),
                        as.numeric(transitivity(x)),
                        as.numeric(average.path.length(x, unconnected=FALSE)), # important
                        as.numeric(mean(graph.strength(x))))
  
                        colnames(graph.measure)<-c("nodes","edges","diameter",
                              "density","reciprocity","transitivity",
                              "average path length","average weighted degree")
                        rownames(graph.measure)<-"x"
                        list(graph.measure)
}
```

### use functions *node.centrality* and *network.measure* to show node-level and network-level measures
#### using these two functions
```{r}
require(network)
one_mode_mat<- as.matrix(read.csv("one_mode_ad_mat.csv", row.names=1))
# use network format to use node.centrality function
one_mode_net=network(one_mode_mat)
node.centrality(one_mode_net)
# use igraph format within network.measure funciton
network.measure(one_mode_graph)
```

#### grouping
sometimes, we need to grouping in SNA, for example, show different groups with different colors. here I show an example using dplyr to assign individuals to groups and then assign different colors to the groups. I use igraph here to plot socio-graph here.
```{r}
require(dplyr)
require(igraph)

sna_test=read.csv("sna_test.csv", head=TRUE)
sna_test1<-sna_test %>% mutate (group=ifelse(id1 %in% c("A","C","D"), "group 1",
                                  ifelse(id1 %in% c("B","E"),"group 2","group 3")))
sna_test2 <- sna_test1 %>% mutate(color=ifelse(group=="group 1","darkolivegreen4",
                                        ifelse(group=="group 2","magenta3","peachpuff4")))

sna_edges<-sna_test2 %>% select(id1,id2)

sna_graph <- graph_from_data_frame(sna_edges, directed = TRUE)

plot(sna_graph,
     vertex.label = sna_graph$id1,
     vertex.color = sna_test2$color, 
     vertex.frame.color = "gray", 
     vertex.label.color = "black", 
     vertex.label.cex = 0.8,
     edge.arrow.size = 0.5)
```

### summary
- There are basically two formats for save network data: adjacency matrix (wide-format) and edgelist (long-format). We can use reshape2 package (http://seananderson.ca/2013/10/19/reshape.html) to convert them mutually. 
- We can use *sna* or *igraph* packages to plot sociographs and write functions to calculate node-level and network-level measures, using sna and igraph format respectively. Note: sna package uses *network* format; *igraph* package uses *igraph* format.
- It is useful to master both packages, converting two foramts mutually and applying them to achieve different purposes.
